Simulation Tools for Small Area Estimation: Introducing the \proglang{R}-package \proglang{saeSim}

### 2014-10-10 saeSim.Rnw
saeSim paper update

@@

% Loading (and installing) necessary packages.
<<echo=FALSE>>=
if(!suppressPackageStartupMessages(require(saeSim))) {
  if(!require(devtools)) {
    install.packages("devtools", repos = "http://cran.at.r-project.org/", dependencies = TRUE)
    library(devtools)
  }

@@

\section{Introduction}
A brief overview of what the reader can expect of this article. What is the aim of this article and where can it be located in science. Discussion of the reproducible research movement and where the package may be beneficial. Packages like Sweave, knitr, R-markdown. How can simulation studies (the \proglang{R}-code) be published? As script? As stand-alone-package? How can ideas for specific data generation tools be shared amongst the small area community and thus be more transparent: saeSim (far-fetched but anyway).

%Graphics should be inserted as PostScript files using the
%package {\tt psfig} or {\tt graphicx}. Bibliographic references
%should be given using {\tt bibtex} in the {\em author-date\/}
%style.

\section{Simulation studies in small area estimation}
The role of simulation studies in the field. Separation into model-based and design-based simulation studies. How does \proglang{saeSim} address these 2 perspectives? Identifying steps/phases/components of simulation studies (data generation, sampling, computation).

\section{saeSim: A simulation framework}
Introducing a simulation study as a data manipulation process. From data generation to estimating models. How can this process be mapped into the \proglang{R}-language. Present the simulation phases as flow-diagram and map the function names to the problem domain. Defining interfaces between phases: \proglang{data.frame} in, \proglang{data.frame} out. What is the difference between a design-based (fixed-population) and model-based (random-population) approach.

@@

@@

\subsection{Data generation}
This is not a data generation tool. However, it supports some useful functions to add random variates to the data. Definition of outliers, see Rao (?) -- how is it addressed in the package. More than linear models, define any response with:
<<eval=FALSE>>=
setup %>% sim_resp_eq(y = g(2 * x) + e)
@

\subsection{Drawing samples}
Is present in unit-level simulation studies. Draw with simple random sampling from the whole population or within cluster/domains. Specify the sample size as integer or fraction and add weights if necessary. Wrappers around \proglang{dplyr::sample\_n} and \proglang{dplyr::sample\_frac}.

\subsection{Adding customized computations}
Adding user specified functions to the simulation process is what separates this package from a mere data generation tool. The interface is simple: Add functions with one argument which is the data at that moment and which return the modified data. I am not sure if this and the sampling section is necessary -- maybe it should just be summarized in the overview section...

\subsection{Separation of designing and running simulations}
First develop a scenario outside any looping structures, then run it R-times:

@@

@@

%\bibliographystyle{plainat}
\bibliography{guidelinesAJS}

>>>>

\usepackage{wrapfig}
\usepackage{graphicx}

@@

% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
<<echo=FALSE>>=
if(!suppressPackageStartupMessages(require(saeSim))) {
  if(!require(devtools)) {
    install.packages("devtools", repos = "http://cran.at.r-project.org/", dependencies = TRUE)
    library(devtools)
  }

@@

\section{Introduction}
Reproducible Research has become a widely discussed topic inside science and also the field of statistics. Thanks to the many mostly open-source tools like the \proglang{R}-language and \LaTeX, and also packages like knitr and Sweave, the integration of source code and text is possible and as a problem solved. In that sense the demand for Literate Programming can be if wanted incorporated in the work flow of a scientist. Not only are tools available to make research reproducible, also the demand of making the analysis of articles reproducible is rising. What this means is, that the source-code and data is published alongside an article. However, the requirements in style and clarity of source code are different from the written words in the article itself. This demand \textit{human readable source-code} has already been expressed by Knuth (1984): 
\begin{quote}
\textit{Instead of imagining that our main task is to instruct a computer what to do, let us concentrate rather on explaining to human beings what we want a computer to do.}
\end{quote}
Small Area Estimation is a growing field inside the field of statistics, where simulation studies play an important role. New statistical models are applied in model-based and design-based simulation studies. Considering the demands in reproducibility we want to propose a framework for simulation studies. This framework addresses three demands. First, making tools for data generation available and reusable. Second, unify the process behind simulation studies inside the field of Small Area Estimation. And third, making source-code of simulation studies available, such that it supports the conducted research in a transparent manner.
In this article we want to introduce a new package for the \proglang{R}-language addressing these demands. We will show the importance of simulation studies in the field. Introduce the simulation framework. And demonstrate how to use the package to map the simulation as a process to \proglang{R}.

\section{Small Area Estimation}
What is small area estimation? 2 Models to introduce. FH + BHF. Why are simulation studies important in the field.

\section{A simulation framework}

%% flow-diagram
\begin{wrapfigure}{R}{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\textwidth]{flowdiagram}
\end{center}
\caption{\label{fig:flowdiagram}Process of simulation.} 
\end{wrapfigure}

Design- vs. model-based simulation, we do not promote one or the other. 
Figure \ref{fig:flowdiagram} illustrates the steps in a simulation. Based on a data table, which may be a real or synthetic population or a set of ID variables representing the hierarchy of the data, you start a simulation.

Introducing a simulation study as a data manipulation process. From data generation to estimating models. How can this process be mapped into the \proglang{R}-language. Present the simulation phases as flow-diagram and map the function names to the problem domain. Defining interfaces between phases: \proglang{data.frame} in, \proglang{data.frame} out. What is the difference between a design-based (fixed-population) and model-based (random-population) approach.

@@

asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd asd

@@

First develop a scenario outside any looping structures, then run it R-times:

@@

\section{Data generation}
This is not a data generation tool. However, it supports some useful functions to add random variates to the data. Definition of outliers, see Rao (?) -- how is it addressed in the package. More than linear models, define any response with:
<<eval=FALSE>>=
setup %>% sim_resp_eq(y = g(2 * x) + e)
@

\section{Adding steps to a simulation}
Is present in unit-level simulation studies. Draw with simple random sampling from the whole population or within cluster/domains. Specify the sample size as integer or fraction and add weights if necessary. Wrappers around \proglang{dplyr::sample\_n} and \proglang{dplyr::sample\_frac}.
Adding user specified functions to the simulation process is what separates this package from a mere data generation tool. The interface is simple: Add functions with one argument which is the data at that moment and which return the modified data. I am not sure if this and the sampling section is necessary -- maybe it should just be summarized in the overview section...

\section{Case study}

@@

%\bibliographystyle{plainat}
\bibliography{saeSim}

### 2014-10-10 saeSim.tex
saeSim paper update

% Loading (and installing) necessary packages.

\section{Introduction}
A brief overview of what the reader can expect of this article. What is the aim of this article and where can it be located in science. Discussion of the reproducible research movement and where the package may be beneficial. Packages like Sweave, knitr, R-markdown. How can simulation studies (the \proglang{R}-code) be published? As script? As stand-alone-package? How can ideas for specific data generation tools be shared amongst the small area community and thus be more transparent: saeSim (far-fetched but anyway).

%Graphics should be inserted as PostScript files using the
%package {\tt psfig} or {\tt graphicx}. Bibliographic references
%should be given using {\tt bibtex} in the {\em author-date\/}
%style.

\section{Simulation studies in small area estimation}
The role of simulation studies in the field. Separation into model-based and design-based simulation studies. How does \proglang{saeSim} address these 2 perspectives? Identifying steps/phases/components of simulation studies (data generation, sampling, computation).

\section{saeSim: A simulation framework}
Introducing a simulation study as a data manipulation process. From data generation to estimating models. How can this process be mapped into the \proglang{R}-language. Present the simulation phases as flow-diagram and map the function names to the problem domain. Defining interfaces between phases: \proglang{data.frame} in, \proglang{data.frame} out. What is the difference between a design-based (fixed-population) and model-based (random-population) approach.

@@

\subsection{Data generation}
This is not a data generation tool. However, it supports some useful functions to add random variates to the data. Definition of outliers, see Rao (?) -- how is it addressed in the package. More than linear models, define any response with:
\begin{Schunk}
\begin{Sinput}
> setup %>% sim_resp_eq(y = g(2 * x) + e)
\end{Sinput}
\end{Schunk}

\subsection{Drawing samples}
Is present in unit-level simulation studies. Draw with simple random sampling from the whole population or within cluster/domains. Specify the sample size as integer or fraction and add weights if necessary. Wrappers around \proglang{dplyr::sample\_n} and \proglang{dplyr::sample\_frac}.

\subsection{Adding customized computations}
Adding user specified functions to the simulation process is what separates this package from a mere data generation tool. The interface is simple: Add functions with one argument which is the data at that moment and which return the modified data. I am not sure if this and the sampling section is necessary -- maybe it should just be summarized in the overview section...

\subsection{Separation of designing and running simulations}
First develop a scenario outside any looping structures, then run it R-times:

@@

\begin{Schunk}
\begin{Sinput}
> simulationResults <- simSetup %>% sim(R = 500, parallel = TRUE, mc.cores = 8)
\end{Sinput}
\end{Schunk}

>>>>

% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.

\section{Introduction}
Reproducible Research has become a widely discussed topic inside science and also the field of statistics. Thanks to the many mostly open-source tools like the \proglang{R}-language and \LaTeX, and also packages like knitr and Sweave, the integration of source code and text is possible and as a problem solved. In that sense the demand for Literate Programming can be if wanted incorporated in the work flow of a scientist. Not only are tools available to make research reproducible, also the demand of making the analysis of articles reproducible is rising. What this means is, that the source-code and data is published alongside an article. However, the requirements in style and clarity of source code are different from the written words in the article itself. This demand \textit{human readable source-code} has already been expressed by Knuth (1984): 
\begin{quote}
\textit{Instead of imagining that our main task is to instruct a computer what to do, let us concentrate rather on explaining to human beings what we want a computer to do.}
\end{quote}
Small Area Estimation is a growing field inside the field of statistics, where simulation studies play an important role. New statistical models are applied in model-based and design-based simulation studies. Considering the demands in reproducibility we want to propose a framework for simulation studies. This framework addresses three demands. First, making tools for data generation available and reusable. Second, unify the process behind simulation studies inside the field of Small Area Estimation. And third, making source-code of simulation studies available, such that it supports the conducted research in a transparent manner.
In this article we want to introduce a new package for the \proglang{R}-language addressing these demands. We will show the importance of simulation studies in the field. Introduce the simulation framework. And demonstrate how to use the package to map the simulation as a process to \proglang{R}.

\section{Small Area Estimation}
What is small area estimation? 2 Models to introduce. FH + BHF. Why are simulation studies important in the field.

\section{A simulation framework}

%% flow-diagram
\begin{wrapfigure}{R}{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\textwidth]{flowdiagram}
\end{center}
\caption{\label{fig:flowdiagram}Process of simulation.} 
\end{wrapfigure}

Design- vs. model-based simulation, we do not promote one or the other. 
Figure \ref{fig:flowdiagram} illustrates the steps in a simulation. Based on a data table, which may be a real or synthetic population or a set of ID variables representing the hierarchy of the data, you start a simulation.

Introducing a simulation study as a data manipulation process. From data generation to estimating models. How can this process be mapped into the \proglang{R}-language. Present the simulation phases as flow-diagram and map the function names to the problem domain. Defining interfaces between phases: \proglang{data.frame} in, \proglang{data.frame} out. What is the difference between a design-based (fixed-population) and model-based (random-population) approach.

@@

First develop a scenario outside any looping structures, then run it R-times:

@@

\begin{Schunk}
\begin{Sinput}
> simulationResults <- simSetup %>% sim(R = 500, parallel = TRUE)
\end{Sinput}
\end{Schunk}

\section{Data generation}
This is not a data generation tool. However, it supports some useful functions to add random variates to the data. Definition of outliers, see Rao (?) -- how is it addressed in the package. More than linear models, define any response with:
\begin{Schunk}
\begin{Sinput}
> setup %>% sim_resp_eq(y = g(2 * x) + e)
\end{Sinput}
\end{Schunk}

\section{Adding steps to a simulation}
Is present in unit-level simulation studies. Draw with simple random sampling from the whole population or within cluster/domains. Specify the sample size as integer or fraction and add weights if necessary. Wrappers around \proglang{dplyr::sample\_n} and \proglang{dplyr::sample\_frac}.
Adding user specified functions to the simulation process is what separates this package from a mere data generation tool. The interface is simple: Add functions with one argument which is the data at that moment and which return the modified data. I am not sure if this and the sampling section is necessary -- maybe it should just be summarized in the overview section...

\section{Case study}
